\chapter*{Acknowledgments}


First and foremost, I would like to thank my parents. To my mom, for always answering the phone, even if that means stopping on the side of the road in the middle of a bike ride, and for enduring the frequent sirens that interrupt our calls. To my dad, for his quiet encouragement and belief, and for keeping me on my toes with our weekly New York Times history quiz competition. Without you both, this thesis would not have been possible. 

I would like to thank my brother and best friend Jake, for his unconventional and probing questions that make me think deeply about the world and my decisions, for his willingness to go on any adventure, and for always inspiring me by example to push my mental and physical limits.

I would like to thank Mark Leininger for his unwavering love and support, and for tolerating my habit of saying ``I'm just going to do five more minutes of work'' when we both know that isn't true. I would also like to thank the newest member of our family, Boh, for always greeting me with excitement when I walk through the door. 

Thank you to Jen Stephenson, my first friend in Baltimore, for always being up for a run and chat. To Ethan Gonazales, Jeremy Ardanuy, Anne Rosenthal and Johan Fagerberg for many cozy dinners, sports watching parties, Baltimore outings, and conversations about my research.

Thank you to David Sheehy, Kannan Mackinam, Jessica Hankinson, Caryn Just, Amelia Tenberg, Maggie Arnold, Jeff Burger and Katja Dunlap for starting off many an early morning with a run that doubles as a vent session, and for never letting my ego get too big. 

Thank you to the former residents at 2025 Pratt: Roxy Alei, Joyce Lin and Alida Voet, for being fantastic housemates, friends, and occasional research collaborators, and for being my allies in the never ending-battle against kitchen mice and fruit flies. Thank you to Carolina Noguiera and Amanda Hixenbaugh for the monthly dinner dates, hot tea, and encouragement. 

Thank you to the Harvard skiing (extended) family: Hannah Hardenbergh, Nina Armstrong, Clara Cousins, Tegan Thorley, Annabel Hagen, Maile Sapp, Rachel Hampton, Charlotte Cole, Liz Roux, Jacob Meyerson, Will Rhatigan, Mack Fitzpatrick, Liam Mcinroy, Everett Sapp, James Kitch, Ian Meyer, Chris Calahan, and J\o{}rgen Grav for keeping in touch and providing support from afar. Thank you to the blockmates: Janet Chen, Will Li, Katie Berry, and Aidan Langston for helping me reflect each month through our newsletter. 

Thank you to Lauren Jacobs and Daniel Perkins for 15 years of friendship. 

Thank you to my classmates in the Biostatistics PhD program, especially Kinnary Shah, Charlotte Clapham, Marina Hernandez, Andrew Chin, Joe Sartini, Cindy Fang, Sowyma Parthiban, Emily Norton, and Grace Ringlein for moral support throughout the PhD process. 

Thank you to members past and present of the Wearable and Implantable Technology (WIT) group, including but not limited to Dr. Vadim Zippunikov, Dr. Xinkai Zhou, Yu Lu, Angela Zhao, and Sam Fansler, for making WIT a dynamic and inclusive research environment. 

Thank you to my cardiac surgery collaborators: Dr. Lee Goeddel, Dr. Nauder Faraday, and Dr. Charles Brown for sharing your data with us and teaching me the ingredients of a fruitful collaboration.

Thank you to Dr. Marie Diener West, Dr. Karen Bandeen Roche, Dr. Daniel Obeng, Dr. John McGready, and Dr. Margaret Taub for mentorship and role modeling in the teaching realm. I hope to one day be as articulate as each of you! 

Thank you to Mary Joy Argo, for reminding me to register for classes and for tireless behind-the-scenes work scheduling exams and defenses. 

Thank you to Dr. John Muschelli, for your invaluable mentorship and advising about school and beyond. I wouldn't have accomplished half as much without the computational skills I learned from working with you. Your open door policy, your willingness to sit down and start working on a problem instead of speaking about it in the abstract, your strategic thinking, and your encouragement to think about career paths has been incredibly helpful. I hope you'll still help me debug (on occasion) in the future. 

Thank you to Dr. Ciprian Crainiceanu for being the best advisor I could have hoped for. Your advising on my life, career and research has been critical, and you struck the perfect balance of giving me guidance and giving me freedom to explore my own ideas. Your prioritization of work-life balance both in your own life and for your students allowed me to thrive and to truly enjoy my time as a PhD student. 



